---
title: "Untitled"
format: html
editor: visual
---

## Part 2: Tidying data in NHANES data

The [**N**ational **H**ealth **A**nd **N**utrition **E**xamination **S**urvey](https://www.cdc.gov/nchs/nhanes/index.html), or NHANES, assess health/nutritional status of adults/children in the United States. It surveys the following types of questions:

-   Demographic (Age, Race, Gender, many more...)

-   Socioeconomic (Marriage Status, Household Income, Education)

-   Dietary (Foods consumed, dietary supplements)

-   Health (Body Mass Index, Sleep Trouble, Depression)

Let's load the data in and take a look.

```{r}
nhanes = read.csv("NHANES.csv")
```

```{r}
Vieww(nhanes)
```

We would like to answer the following question: What is the relationship between pulse and BMI categories for \_\_\_ vs \_\_\_\_?

```{r}

```

First, some data cleaning:

If you looko at the `ID` column of `nhanes` Dataframe, you will see that it is numeric value, but we won't be using it as a number value - they are categorical information representing each patient. Use the `as.character()` function to convert the `ID` column to be character type.

```{r}

```

Take a look at the `Pulse` column. Notice that it has a lot of missing value as `NA`. Remove all rows of `nhanes` with `NA` in the `Pulse` column.

```{r}

```

BMI categorizations are only useful for adults. Remove all rows of `nhanes` with `Age` less than 20.

```{r}

```

```         
```
