# Fundamentals Exercises

## Part 1: Vectors

How do you subset the following vector to the first three elements?

```{r}
measurements = c(2, 4, -1, -3, 2, -1, 10)
```

```{r}

```

How do you subset the original vector so that it only has negative values?

```{r}

```

How do you subset the following vector so that it has no NA values?

```{r}
vec_with_NA = c(2, 4, NA, NA, 3, NA)
```

```{r}

```

Consider the following logical vector `some_logicals`. Coerce the data type from Logical vector -\> Numeric vector -\> Character vector in two steps. Check that you are doing this correctly along the way by using the `class()` function, or `is.numeric()` and `is.character()`, on the converted data.

```{r}
some_logicals = c(TRUE, TRUE, TRUE, FALSE, TRUE, FALSE)
```

```{r}

```

## Part 2: Lists

Consider the following lists with names.

```{r}
patient = list(
  name = " ", 
  age =  34, 
  pronouns = c("he", "him", "/", "they", "them"),
  vaccines = c("hep-B", "chickenpox", "HPV"),
  visits = NA
)

visit1 = list(
  symptoms = c("runny nose", "sore throat", "frustration"),
  prescription = "recommended time off from work, rest.",
  date = "1/1/2000"
)

visit2 = list(
  symptoms = c("fainted", "pale complexion"),
  prescription = "drink water and take time off work.",
  date = "1/1/2001"
)
```

Access the first list element of `patient` via double brackets [[ ]] or \$ and modify it to a value of your choice.

```{r}

```

Access the list element "pronouns" of `patient` via double bracket [[ ]] or \$ and modify its value so that it doesn't contain the "/" element. You can do this two ways: assign an entire new vector to `patient$pronouns`. Or, update `patient$pronouns` via vector subsetting.

```{r}

```

Now, we create a new list `all_visits` with elements `visit1` and `visit2`. Yes, we're making lists within lists!

```{r}
all_visits = list(visit1, visit2)
```

Take a look at `all_visits` yourself.

```{r}

```

Suppose you want to use `all_visits` to access visit 1's symptoms. You would continue the double brackets [[ ]] or \$ notation: `all_visits[[1]]` returns a list, so we access the first element of *that* list via `all_visits[[1]][[1]]`.

```{r}
all_visits[[1]][[1]]

#or

all_visits[[1]][["symptoms"]]

#or

all_visits[[1]]$symptoms
```

How would you use `all_visits` to access visit 2's prescription?

```{r}

```

How would you use `all_visits` to access visit 2's symptom element "pale complexion"? Remember, once you access a vector, you would go back to the single bracket [ ] to access its elements.

```{r}

```

Finally, assign `all_visits` to `patient`'s visits (ie. replace the `NA` value of `patient$visit` with `all_visits`).

```{r}

```

Take a look at `patient` now. Could you comfortably access elements in this list? :

```{r}

```

## Part 3: Dataframes (Lists)

A dataframe is just a named list of vectors of same length with **attributes** of (column) `names` and `row.names`.

```{r}
library(palmerpenguins)
head(penguins)
```

Access the `body_mass_g` column vector of `penguins` via the double bracket [[ ]] or \$ notation, treating `penguins` like a list, and compute the mean. Remember to use na.rm = TRUE to remove any `NA` values: `mean(x, na.rm = TRUE)`

```{r}

```

Create a new dataframe `penguins_clean`, which has no `NA` values in the `body_mass_g` column. You need to filter out rows that have `NA`s in the column `bill_length_mm`: (Hint: look at classwork example.)

```{r}

```

Now, subset `penguins_clean` to each of the three species and compute their respective mean value of `body_mass_g`. Because you already got rid of `NA`s in `body_mass_g`, you can just use `mean(x)` without the extra option. How do they compare?

```{r}

```

Finally, make a box plot of `species` (x-axis) vs. `body_mass_g` (y-axis) via `penguins_clean` dataframe. I'll get you started...

```{r}
#ggplot(penguins_clean) + aes(x = , y = ) + geom_boxplot()
```


## Feedback!

How many hours did you spend on this exercise?

```{r}
time_spent = 0 
```

If you worked with other peers, write their names down in the following character vector: each element is one person's name.

```{r}
peers = c("myself")
```