---
title: "W2: Data Cleaning 1"
format: 
  revealjs:
    smaller: true
    scrollable: true
    echo: true
output-location: fragment
---

```{r, echo=F, message=F, warning=F}
library(tidyverse)
library(palmerpenguins)
```

## Review: When do you refer columns with a `$`?

. . .

Tidyverse Dataframe functions, such as `select()`, `filter()`, `mutate()`, `group_by()`, `summarize()` allows you to refer columns without `$`:

```         
select(penguins, species, island)

filter(penguins, species == "Adelie")

mutate(penguins, new_col = log(body_mass_g))

penguins |> group_by(island) |> summarize(avg_body_mass_g = mean(body_mass_g))
```

. . .

All other functions and operations require `$` as a reference:

```         
mean(penguins$body_mass_g, na.rm = TRUE)

table(penguins$island)

penguins$island[penguins$island == "Torgersen"]
```

## Recoding Data / Conditionals

. . .

Suppose that you have a column in your dataframe that needs to be **recoded**.

. . .

```{r}
simple_df = data.frame(grade = c(90, 78, 95, 74, 56, 81, 102),
                       status = c("case", " ", "Control", "control", "Control", "Case", "case"))
simple_df
```

. . .

Since a dataframe's column, when selected via `$`, is a vector, let's start talking about **recoding vectors**.

## Common recoding scenarios:

1.  **If**: "If elements of the vector meets *condition*, then they are assigned *value*."

. . .

2.  **If-else**: "If elements of the vector meets *condition*, then they are assigned *value X*. Otherwise, they are assigned *value Y*."

. . .

3.  **If-else_if-else**: "If elements of the vector meets *condition A*, then they are assigned *value X*. Else, if the elements of the vector meets *condition B*, they are assigned *value Y*. Otherwise, they are assigned *value Z*."

## Recoding a vector

```{r}
grade = c(90, 78, 95, 74, 56, 81, 102)
```

. . .

1.  If `grade` is greater than 100, then they are assigned to 100.

. . .

```{r}
grade1 = grade

grade1[grade1 > 100] = 100
grade1
```

. . .

2.  If `grade` is greater than 60, then it is assigned to "Pass". Otherwise, then it is assigned to "Fail".

. . .

```{r}
grade2 = if_else(grade > 60, "Pass", "Fail")
grade2
```

## Recoding a vector

3.  If `grade` is greater than 90, then it is assigned to `"A"`. Else, if `grade` is greater than 80, then it is assigned to `"B"`. etc.\
    Otherwise, they are assigned to `"F"`.

```{r}
grade3 = case_when(grade >= 90 ~ "A",
                   grade >= 80 ~ "B",
                   grade >= 70 ~ "C", 
                   grade >= 60 ~ "D",
                   .default = "F")
grade3
```

## Recoding a dataframe column

```{r}
simple_df
```

. . .

1.  If

    ```{r}
    simple_df1 = simple_df

    simple_df1$grade[simple_df1$grade > 100] = 100
    simple_df1
    ```

. . .

2.  If-else

```{r}
simple_df2 = simple_df
simple_df2$grade = if_else(simple_df2$grade > 60, "Pass", "Fail")
simple_df2
```

. . .

or

```{r}
simple_df2 = mutate(simple_df, grade = if_else(grade > 60, "Pass", "Fail"))
```

. . .

## Recoding a dataframe column

3.  If-else_if-else

```{r}
simple_df3 = simple_df

simple_df3$grade = case_when(simple_df3$grade >= 90 ~ "A",
                             simple_df3$grade >= 80 ~ "B",
                             simple_df3$grade >= 70 ~ "C", 
                             simple_df3$grade >= 60 ~ "D",
                             .default = "F")
simple_df3
```

. . .

or

```{r}
simple_df3 = mutate(simple_df, grade = case_when(grade >= 90 ~ "A",
                                                 grade >= 80 ~ "B",
                                                 grade >= 70 ~ "C", 
                                                 grade >= 60 ~ "D",
                                                 .default = "F"))
```

## More Alternatives of a simple "if" statement

```{r}
simple_df$status= if_else(simple_df$status == "case", "Case", simple_df$status)

simple_df = mutate(simple_df, status = if_else(status == "case", "Case", status))

simple_df$status= case_when(simple_df$status == "case"~ "Case", 
                            .default = simple_df$status) 

simple_df$status= case_match(simple_df$status,
                             "case" ~ "Case", 
                             .default = simple_df$status) 
```

<https://dplyr.tidyverse.org/reference/case_match.html#connection-to-case-when->

## Conditionals

The 3 common scenarios we looked at for recoding data is closely tied to the concept of **conditionals** in programming: *given certain conditions, you run a specific code chunk.*

. . .

A conditional statement checks whether an expression has a value of `TRUE`. If so, then it runs an indented code check enclosed by brackets.

Notice that this expression is a single logical value, *not* a logical vector.

. . .

```         
if(expression_is_TRUE) {
  #code goes here
}
```

. . .

```         
if(expression_A_is_TRUE) {
  #code goes here
}else if(expression_B_is_TRUE) {
  #other code goes here
}else {
  #some other code goes here
}
```

. . .

```{r}
subtype = "lung"

if(subtype == "lung") {
  treatment_code = 123
}else if(subtype == "prostate") {
  treatment_code = 234
}else if(subtype == "kidney") {
  treatment_code = "345"
}else if(subtype == "bone") {
  treatment_code = "4757"
}else {
  treatment_code = NA
}

treatment_code
```

## More examples

```{r}
year = rnorm(100, mean=1990, sd=4)
```

```{r}
if(is.numeric(year)) {
  df = data.frame(year = year)
  ggplot(df) + aes(x = year) + geom_histogram()
}else {
  warning("year is not numeric for plotting!")
}
```
